---
title: "ESM 206 Lab 5"
author: "Oswaldo Felix"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Lab 5 objectives

- Learn to parse dates with 'lubridate'
- Get counts observations with count()
- Uncount() rows
- One and two-sample t-test
- Create a heatmap with geom_tile()

```{r}
# Read in data and attach packages
library(tidyverse)
library(here)
library(janitor)

lobster_abundance <- read_csv(here::here("data", "lobster_abundance.csv"), 
                              na = "-99999") %>% 
  clean_names()
  
```

Use tidyr::uncount() function to convert our lobster data from frequency format to case format

```{r}
lobster_tidy <- lobster_abundance %>% 
  uncount(lobster_count)
```

#### Exploratory data visualization

We're only going to consider 'site' as our variable of interest

```{r}

ggplot(lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site,
                  alpha = 0.5),
                  width = 0.2)

# Histogram 
ggplot(lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free")

# qq-plot

ggplot(lobster_tidy, aes(sample = size_mm)) +
  geom_qq() +
  facet_wrap(~site)
```

#### Convert the 'date' colum to class 'Date'

We'll use the 'lubridate' package to convert to Date format, and then to help us easily parse month & year

```{r}

lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date)
    )
```

Now let's parse year and month using lubridate::month() and lubridate::year():

```{r}

lobster_parse_date <- lobster_date %>% 
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE),
    obs_year = lubridate::year(date_new)
  )
```

Now let's find counts of observed lobsters based on different hierarchical groupings:

First let's count lobsters by year and month
```{r}
lobster_ym <- lobster_parse_date %>% 
  count(obs_year, obs_month)

#lobster_ym

lobster_y <- lobster_parse_date %>% 
  count(obs_year)

#lobster_y

lobster_s <- lobster_parse_date %>% 
  count(site)

#lobster_s
```

If we want to create a summary table that contains statistics OTHER than counts by group, it's easier to use group_by() + n()

```{r}
lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    lobster_number = n()
  )

lobster_summary
```

#### Find confidence intervals

Use t.test() function to find confidence intervals (for one sample) and perform t-tests to compare means of two samples (...this will be covered conceptually in lectures Week 6)

```{r}
ivee_lobster <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

t.test(ivee_lobster)
#ivee_lobster
```

#### Two-sample t-test to compare means

```{r}

```

